---
title: "Epidome sequence analysis"
output:
  html_notebook:
    toc: yes
author: "Thor Bech Johannesen, Statens Serum Institut"
---


## **Classification of dada2 generated ASVs, creation of barplots, PCA plots and more**

### Load functions

```{r}
source("https://raw.githubusercontent.com/ssi-dk/epidome/master/scripts/epidome_functions.R")
```
#### <br></br>
#### Read in data. Required data is:
###### Table containing frequencies of STs and amplicons (found in github repo)
###### Sequence count tables in dada2 format created by running ASV_blast_classification.py on dada2 ASV table output
###### OPTIONAL: metadata table. Row names must match column names in dada output.
```{r}
ST_amplicon_table = read.table("https://raw.githubusercontent.com/ssi-dk/epidome/master/DB/epidome_ST_amplicon_frequencies.txt",sep = "\t")
epi01_table = read.table("https://raw.githubusercontent.com/ssi-dk/epidome/master/example_data/190920_run1_G216_seqtab_from_dada2.csv.classified.csv",sep = ";",header=TRUE,row.names=1)
epi02_table = read.table("https://raw.githubusercontent.com/ssi-dk/epidome/master/example_data/190920_run1_yycH_seqtab_from_dada2.csv.classified.csv",sep = ";",header=TRUE,row.names=1)
metadata_table = read.table("https://raw.githubusercontent.com/ssi-dk/epidome/master/example_data/article_metadata.txt",header=TRUE,row.names=1)
```
#### <br></br>
#### If ASV data is spread over multiple dada runs they can be combined like this
```{r}
epi01_mock_table = read.table("https://raw.githubusercontent.com/ssi-dk/epidome/master/example_data/190920_run2_G216_seqtab_from_dada2.csv.classified.csv",sep = ";",header=TRUE,row.names=1)
epi02_mock_table = read.table("https://raw.githubusercontent.com/ssi-dk/epidome/master/example_data/190920_run2_yycH_seqtab_from_dada2.csv.classified.csv",sep = ";",header=TRUE,row.names=1)
epi01_clinical_table = read.table("https://raw.githubusercontent.com/ssi-dk/epidome/master/example_data/190920_run1_G216_seqtab_from_dada2.csv.classified.csv",sep = ";",header=TRUE,row.names=1)
epi02_clinical_table = read.table("https://raw.githubusercontent.com/ssi-dk/epidome/master/example_data/190920_run1_yycH_seqtab_from_dada2.csv.classified.csv",sep = ";",header=TRUE,row.names=1)
epi01_table = combine_ASV_tables(epi01_mock_table,epi01_clinical_table)
epi02_table = combine_ASV_tables(epi02_mock_table,epi02_clinical_table)
```


#### <br></br>
#### Setup object for data handling
##### without metadata
```{r}
epidome_object = setup_epidome_object(epi01_table,epi02_table)
```
#### <br></br>
##### with metadata
```{r}
epidome_object = setup_epidome_object(epi01_table,epi02_table,metadata_table = metadata_table)
epidome_object$p1_table
epidome_object$p2_table
epidome_object$metadata
```
#### <br></br>
#### Check how well the output from the two primer sets align. Color_variable can be left out
```{r}
primer_compare = compare_primer_output(epidome_object,color_variable = "sample.type")
primer_compare$plot
```

##### Combine ASVs assigned to the same taxa
```{r}
eo_ASV_combined = combine_ASVs_epidome(epidome_object)
```
#### <br></br>


##### Remove low count samples from data set. (Only include samples where number of sequences from primer1 >= x and sequences form primer2 >= y)
```{r}
eo_filtered = filter_lowcount_samples_epidome(eo_ASV_combined,500,500)
```
#### <br></br>

#####  --------------------------------------------------------
##### Assign ASVs to STs on a per sample basis and make barplot
```{r}
count_table = classify_epidome(eo_filtered,ST_amplicon_table)
p = make_barplot_epidome(count_table,reorder=FALSE,normalize=TRUE)
p
```


#### <br></br>

#### Data stratification based on metadata variables (requires metadata)
```{r}
eo_clinical = prune_by_variable_epidome(epidome_object,"sample.type",c("Clinical"))
eo_mock = prune_by_variable_epidome(epidome_object,"sample.type",c("Mock community"))
```
#### <br></br>

#####  ------------------------------------------------------------------------------------------------
#### make PCA plots and color by given variable (requires metadata)
##### Normalize data to percent
```{r}
epidome_object_clinical_norm = normalize_epidome_object(eo_clinical) ### Normalize counts to percent
```

##### Color by patient
```{r}
PCA_patient_colored = plot_PCA_epidome(epidome_object_clinical_norm,color_variable = "patient.ID",colors = c(),plot_ellipse = FALSE)
PCA_patient_colored + ggtitle("PCA plot of nose and skin samples colored by subject")

```

#### <br></br>

##### Color by sampling site
```{r}
PCA_sample_site_colored = plot_PCA_epidome(epidome_object_clinical_norm,color_variable = "sample.site",colors = c("Red","Blue"),plot_ellipse = TRUE)
PCA_sample_site_colored + ggtitle("PCA plot of nose and skin samples colored by sampling site")

```
#### <br></br>

#### General data manipulation
##### Remove low count samples
```{r}
eo_filter_lowcount = filter_lowcount_samples_epidome(epidome_object,p1_threshold = 500,p2_threshold = 500)
```
##### Remove low count ASVs
```{r}
eo_filter_ASVs = epidome_filtered_ASVs = filter_lowcount_ASVs_epidome(epidome_object,percent_threshold = 1)
```
##### Normalize data to percent
```{r}
epidome_object_normalized = normalize_epidome_object(epidome_object)
```

##### Combine ASVs that have been classified as the same sequence
```{r}
epidome_object_ASV_combined = combine_ASVs_epidome(epidome_object)
```
##### Make epidome object only including isolates were the variable in "variable_name" is equal to one found in "variable_values"
```{r}
epidome_object_clinical = prune_by_variable_epidome(epidome_object,variable_name = "sample.type",variable_values = c("Clinical"))
```

